---
title: "Introduction to `ExtendedRtIrtModeling.jl` through `JuliaConnectoR`"
author: "JW Tsai"
date: "2024-11-19"
execute:
  cache: false
---

I've just updated my Julia package, `ExtendedRtIrtModeling.jl`, to version 0.2.0. There are a few new features in there that I'll run through in the next few sections.

But that's not all! If you're an R user, I'll introduce you to an R package called `JuliaConnectoR` that lets you run Julia programs in R. It'll bridge the two languages seamlessly.

## Using in Julia

See the Github page.

## Using in R

All you have to do is to install and library the `JuliaConnectoR` as usual, and then you can use the `juliaImport` function to import any Julia package. It seems like the package's version you get depends on which copy version you've had on your computer (**confirmed**). ~~The great thing is, it'll always download the newest version from Github, but not the stable one.~~

```{r}
library(tidyverse)
library(JuliaConnectoR)
```

I've got a toy data set to test, but I'll run through the demo anyway. As you can see from the data below, the data set includes 25 columns: one for ID, 10 for item responses, 10 for (log-)response time, and four for explanatory variables.

```{r}
demo <- read.csv('https://raw.githubusercontent.com/jiewenTsai/ExtendedRtIrtModeling.jl/refs/heads/main/data/demo.csv')
head(demo)
glimpse(demo)
```

Next, let's take a look at how accuracy and speed related to each other, barely using the raw data (mean of y and mean of t) to get a rough idea.

```{r}

tibble(
  accuracy = rowMeans(demo[2:11]),
  speed = -rowMeans(demo[12:21])
) |>
  ggplot(aes(x=accuracy, y=speed)) +
  geom_point() +
  geom_jitter() +
  geom_smooth(method = "loess") 

```

## The modeling part.

Let's follow the example from the Github readme post to show how a basic RT-IRT model works. This package is version 0.2.0.

```{r}

## You have to give a name to the Julia Environment.
ex <- juliaImport("ExtendedRtIrtModeling")
juliaEval('Pkg.status("ExtendedRtIrtModeling")')
```

In the original Github post, the user doesn't have to fill in all five arguments of `InputData` because the values for $\kappa$ and *logT* are automatically calculated by *Y* and *T*. However, to get it working with R, I've come up with another struct called `InputData4R`, which is specifically for R users.

```{r}
## import your data set
Cond = ex$setCond(
  nChain=3, 
  nIter=3000,
  nSubj=300,
  nItem=10,
  nFeat=4
  )
Data = ex$InputData4R(
    Y = as.matrix(demo[2:11]),
    # you must write this line!
    κ = as.matrix(demo[2:11]-0.5),
    T = as.matrix(exp(demo[12:21])),
    logT = as.matrix(demo[12:21]),
    X = as.matrix(demo[22:25])
)

## build a model and sample it!
MCMC = ex$GibbsRtIrt(Cond, Data=Data)
## Notice the `` style for sample!
ex$`sample!`(MCMC)

ex$coef(MCMC)
ex$precis(MCMC)

```

If you want to use the `MCMC` class data in R, you can use the `juliaGet` function to convert Julia's objects to R's. But it'll take a loooooot of time.

```{r}
#| cache: true
time <- Sys.time()
MCMC4R <- juliaGet(MCMC)
Sys.time() - time

ls(MCMC4R)
ls(MCMC4R$Post$mean)
```

```{r}
MCMC4R$Post$ra[,311,] |> 
  as.data.frame() |>
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value") %>%
  ggplot(aes(x = 1:nrow(.), y = value, color = variable, group = variable)) +
  geom_line() +
  labs(x = "Index", y = "Value", title = "b1 MCMC chains")

```

Lastly, let's check out this package.

```{r}
## See the objects in ex.
ls(ex)
```
